Nitrogen's form determines its fate in the North Pacific Subtropical Gyre during PARAGON I

### 2025-05-12 manuscript/manuscript.Rmd
Massively cleaning up manuscript directory

---
title: ""
author:
  - William Kumler
  - Laura T. Carlson
  - Anitra E. Ingalls
output: 
  word_document:
    reference_docx: Frontiers_Template.docx
csl: frontiers-in-marine-science.csl
bibliography:
  - r-packages.bib
---

@@

pdf("thaas/all_cmpd_labels.pdf", width=9, height = 5)
for(cmpd_i in setdiff(unique(thaa_quant$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
  print(cmpd_i)
  gp <- thaa_quant %>%
    filter(compound_name==cmpd_i) %>%
    left_join(metathaa) %>%
    drop_na() %>%
    ggplot() +
    geom_col(aes(x=tripl, y=env_conc_nm, fill=iso_name), 
             position="fill",
             color="black", width=1) +
    facet_nested(depth+startime~amendment+timepoint) +
    ggtitle(cmpd_i)
  print(gp)
}
dev.off()
```

>>>>

---
title: "Nitrogen's form determines its fate in the North Pacific Subtropical Gyre during PARAGON I"
author:
  - William Kumler
  - Laura T. Carlson
  - Anitra E. Ingalls
  - Other SCOPE collaborators
output: word_document
---

@@

pdf("figures/all_cmpd_labels.pdf", width=9, height = 5)
for(cmpd_i in setdiff(unique(thaa_quant$compound_name), c("Citrulline", "Ornithine", "Taurine"))){
  print(cmpd_i)
  gp <- thaa_quant %>%
    filter(compound_name==cmpd_i) %>%
    left_join(metathaa) %>%
    drop_na() %>%
    ggplot() +
    geom_col(aes(x=tripl, y=env_conc_nm, fill=iso_name), 
             position="fill",
             color="black", width=1) +
    facet_nested(depth+startime~amendment+timepoint) +
    ggtitle(cmpd_i)
  print(gp)
}
dev.off()
```

## Supp fig: Fraction of AA as DFAA vs THAA

```{r fraction of aa pool in df vs th}
thaa_pool_sizes <- thaa_quant %>%
  filter(!compound_name%in%c("Citrulline", "Ornithine", "Taurine")) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  filter(str_detect(filename, "Smp.*T0")) %>%
  select(filename, compound_name, env_conc_nm) %>%
  mutate(compound_name=ifelse(compound_name=="Isoleucine", "Leucine", compound_name)) %>%
  summarise(env_conc_nm=sum(env_conc_nm), .by=c(filename, compound_name)) %>%
  mutate(compound_name=ifelse(compound_name=="Leucine", "Leu/Ile", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glu/Gln", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name=="Aspartic acid", "Asp/Asn", compound_name)) %>%
  left_join(metathaa) %>%
  summarise(mean_thaa_conc=mean(env_conc_nm), .by=c(amendment, depth, startime, compound_name))
metab_pool_sizes <- env_concs %>%
  filter(str_detect(filename, "Smp.*T0")) %>%
  mutate(compound_name=str_remove_all(compound_name, "L-")) %>%
  mutate(compound_name=str_remove(compound_name, "/Homoserine")) %>%
  mutate(compound_name=str_remove(compound_name, "/beta-Glutamic acid")) %>%
  mutate(compound_name=ifelse(compound_name=="Glutamine", "Glutamic acid", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name=="Asparagine", "Aspartic acid", compound_name)) %>%
  summarise(conc_nm_in_env=sum(conc_nm_in_env), .by=c(filename, compound_name)) %>%
  mutate(compound_name=ifelse(compound_name=="Glutamic acid", "Glu/Gln", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name=="Aspartic acid", "Asp/Asn", compound_name)) %>%
  mutate(compound_name=ifelse(compound_name=="Leucine/Isoleucine", "Leu/Ile", compound_name)) %>%
  filter(compound_name%in%thaa_pool_sizes$compound_name) %>%
  left_join(parametadata) %>%
  summarise(mean_metab_conc=mean(conc_nm_in_env), .by=c(amendment, depth, startime, compound_name))
# Valine and phenylalanine failed QC
thaa_pool_sizes %>%
  left_join(metab_pool_sizes) %>%
  ggplot() +
  geom_point(aes(x=mean_thaa_conc, y=mean_metab_conc, color=compound_name)) +
  facet_nested(depth+startime~amendment)
thaa_pool_sizes %>%
  left_join(metab_pool_sizes) %>%
  ggplot() +
  geom_point(aes(x=mean_thaa_conc, y=mean_metab_conc)) +
  facet_wrap(~compound_name, scales="free")

thaa_pool_sizes %>%
  left_join(metab_pool_sizes) %>%
  mutate(metab_frac=mean_metab_conc/mean_thaa_conc) %>%
  ggplot() +
  geom_density(aes(x=metab_frac, color=compound_name))
thaa_pool_sizes %>%
  left_join(metab_pool_sizes) %>%
  mutate(metab_frac=mean_metab_conc/mean_thaa_conc) %>%
  ggplot() +
  geom_boxplot(aes(x=compound_name, y=metab_frac)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x=NULL, y='Fraction of AA in "free" pool: [metab] / [THAA]') +
  coord_flip()
```
